Analyzing X-ray images in Python with scikit-image

### 2016-08-19 skimage_tomo.tex
Iteration on performance + doc

\texttt{scikit-image} is only one Python module that can be used for data
processing, among many others. A very popular module is
\texttt{scikit-learn} \citep{Pedregosa2011}, a Python module for machine
learning using \texttt{NumPy} arrays. If local features of an image
(such as local statistics of gray levels, or geometric points of
interest), or features of segmented objects (e.g. geometrical and
intensity characteristics of segmented particles) are extracted with
functions from \texttt{skimage.feature} (see Fig.~\ref{fig:ecosystem}, it
is then possible to use a \emph{classification} algorithm of
\texttt{scikit-learn}, either for labeling pixels (a segmentation task)
or to classify whole images or objects that have already been segmented.
The interoperability between these packages is ensured by the common use
of \texttt{NumPy} arrays, so that the tasks described above are
implemented with a few lines of code.

The modularity of the Scientific Python ecosystem may look
confusing at first sight. Nevertheless, core modules of this ecosystem
are almost perfectly compatible, thanks to the shared use of NumPy arrays
and common development practices (although they are developed in parallel
by different teams).

@@

Tomographic images often suffer from artifacts or poor signal-to-noise
ratio. Several denoising filters are available for image restoration,
from general-purpose filters such as median and bilateral filters, to more
specific filters. Total-variation denoising \citep{Chambolle2004,
Getreuer2012} is well-suited to restore
piecewise-constant images, such as images with a small number of phases
encountered in materials science \citep{Bouttes2015}. Conversely, images
with a fine-grained texture are better preserved with non-local means
denoising \citep{Buades2005} (see Fig.~\ref{fig:tomo_gallery}).

@@

\emma{Juan or Stéfan, can you add something about dask?}

Prototyping the image processing workflow requires several trial and error
executions of scripts containing different operations. The \texttt{joblib}
module provides a caching mechanism that avoids the repetition of function
calls, if their arguments have not changed:
\begin{lstlisting}
>>> from skimage import filters, data
>>> from joblib import Memory
>>> mem = Memory(cachedir='/tmp/joblib')
>>> median = mem.cache(filters.median)
>>> im = data.camera()
>>> filtered_im = median(im, np.ones((3, 3)))
__________________________
[Memory] Calling skimage.filters.rank.generic.median...
median(array([[156, ..., 152],
       ..., 
       [121, ..., 111]], dtype=uint8), array([[ 1.,  1.,  1.],
       [ 1.,  1.,  1.],
       [ 1.,  1.,  1.]]))
__________________________median - 0.0s, 0.0min
>>> filtered_again = median(im, np.ones((3, 3)))
>>> # the above call did not trigger an evaluation
\end{lstlisting}

@@

The success of a software package aimed at end-users is strongly
correlated to the quality of its documentation (find citation).
\texttt{scikit-image} users have access to different kinds of
documentation. Functions of the API (define API here or above) are all
documented using the same standard \citep{Pawlik2015} as other major
Scientific Python packages. The standard include a description of all
input and output variables and their data type, together with
explanations on what the function does and how to use it. Function
documentation is accessed either online, or within the development
environment (IPython, Spyder, Jupyter Notebook...).

@@

In addition, a popular documentation tool is the graphical gallery of
examples (\url{http://scikit-image.org/docs/dev/auto_examples/}), a part of which is displayed in Fig.~\ref{fig:gallery}.
The principle of the gallery is to accelerate the learning curve of image
processing, by showcasing graphical examples of image processing
operations. Such examples are organized as an array of thumbnails with a
short title (see Fig.~\ref{fig:gallery} left). These thumbnails link to
the webpage of the corresponding example, which features a mini-tutorial
on the image processing method, the code needed to run the example and
the figure generated by the example. Since the graphical gallery is an
efficient way to inform users about the features of
\texttt{scikit-image}, every new feature integrated in the package must
include an example for the gallery.

Longer tutorials and a more narrative documentation is available as well
in the User Guide of \texttt{scikit-image}. The User Guide explains in
particular...

@@

Estimating the number of active users of an open-source package is a
difficult task. Statistics of downloads, for example, largely
overestimate the number of active users, all the more if the package is
bundled with others in a software distribution, like Anaconda or Canopy.
A view closer to reality can be obtained by analyzing the statistics of
visits of the online help, available on the project website.  As of the
first half of 2016, 20000 unique visitors per month visit the website of
scikit-image \url{http://scikit-image.org/}. These users come from 80
countries, located almost equally in Europe, America and Asia (Oceania
and Africa representing a much smaller fraction of visitors).

@@

>>>>

\texttt{scikit-image} is only one Python module that can be used for data
processing, among many others. A very popular module is
\texttt{scikit-learn} \citep{Pedregosa2011}, a Python module for machine
learning using \texttt{NumPy} arrays. If local features of an image
(such as local statistics of gray levels, or geometric points of
interest), or features of segmented objects (e.g. geometrical and
intensity characteristics of segmented particles) are extracted with
functions from \texttt{skimage.feature} (see Fig.~\ref{fig:ecosystem}), it
is then possible to use a \emph{classification} algorithm of
\texttt{scikit-learn}, either for labeling pixels (a segmentation task)
or to classify whole images or objects that have already been segmented.
The interoperability between these packages is ensured by the common use
of \texttt{NumPy} arrays, so that the tasks described above are
implemented with a few lines of code.

The modularity of the Scientific Python ecosystem may be
confusing at first sight. Nevertheless, core modules of this ecosystem
are almost perfectly compatible, thanks to the shared use of NumPy arrays
and common development practices (although they are developed in parallel
by different teams). Several integrated distributions, such as Anaconda
or Canopy, bundle together the most common packages, including
\texttt{scikit-image}.

@@

Tomographic images often suffer from artifacts or poor signal-to-noise
ratio. Several denoising filters are available for image restoration,
from general-purpose filters such as median and bilateral filters, to
more specific filters. Total-variation denoising \citep{Chambolle2004,
Getreuer2012} is well-suited to restore piecewise-constant images (see
Fig.~\ref{fig:tomo_gallery}), such as images with a small number of
phases encountered in materials science \citep{Bouttes2015}. Conversely,
images with a fine-grained texture are better preserved with non-local
means denoising \citep{Buades2005} (see Fig.~\ref{fig:tomo_gallery}).

@@

An easy solution consists in dividing an image into smaller images (with
or without overlap, depending on the operation), and to apply the
same operation on the different sub-images, on different cores. Creating overlapping chunks is easy with the dedicated function \texttt{view\_as\_windows} (or \texttt{view\_as\_blocks} for contiguous non-overlapping chunks):
\begin{lstlisting}[language=Python]
>>> from skimage import util, data
>>> im = data.camera()
>>> im.shape
(512, 512)
>>> # chunks of size 120 with 8-pixel overlap
>>> chunks = util.view_as_windows(im, (128, 128), step=120)
>>> chunks.shape
(4, 4, 128, 128)
\end{lstlisting}
The \texttt{joblib} module enables easy parallel processing. Looping over the different blocks, and dispatching the computation over several cores, is realized with the following syntax:  
\begin{lstlisting}
>>> from joblib import Parallel, delayed
>>> filtered_chunks = Parallel(n_jobs=4)(delayed(filters.gaussian)(chunks[i, j]) for i in range(4) for j in range(4))
\end{lstlisting}
\texttt{scikit-image} also offers experimental support for a more integrated parallel processing pipeline, thanks to the \texttt{dask} \citep{dask} module:
\begin{lstlisting}
filtered_im = util.apply_parallel(filters.gaussian, im, depth=8)
\end{lstlisting}
The size of chunks is determined automatically from the number of available cpus, or can be specified by the user.

On a different note, prototyping the image processing workflow requires several trial and error
executions of scripts containing different operations. The \texttt{joblib}
module provides a caching mechanism that avoids the repetition of function
calls, if their arguments have not changed:
\begin{lstlisting}
>>> from skimage import filters, data
>>> from joblib import Memory
>>> mem = Memory(cachedir='/tmp/joblib')
>>> median = mem.cache(filters.median)
>>> im = data.camera()
>>> filtered_im = median(im, np.ones((3, 3)))
__________________________
[Memory] Calling skimage.filters.rank.generic.median...
median(array([[156, ..., 152],
       ..., 
       [121, ..., 111]], dtype=uint8), array([[ 1.,  1.,  1.],
       [ 1.,  1.,  1.],
       [ 1.,  1.,  1.]]))
__________________________median - 0.0s, 0.0min
>>> filtered_again = median(im, np.ones((3, 3)))
>>> # the above call did not trigger an evaluation
\end{lstlisting}

@@

The quality of software documentation is an important component to the
success of a package aimed at scientists.
\texttt{scikit-image} users have access to different kinds of
documentation. Functions of the API are all
documented using the same standard \citep{Pawlik2015} as other major
Scientific Python packages. The standard include a description of all
input and output variables and their data type, together with
explanations on what the function does and how to use it. Function
documentation is accessed either online, or within the development
environment (IPython, Spyder, Jupyter Notebook...).

@@

In addition, a popular documentation tool is the graphical gallery of
examples (\nolinkurl{http://scikit-image.org/docs/dev/auto_examples/}), a part of which is displayed in Fig.~\ref{fig:gallery}.
The principle of the gallery is to accelerate the learning curve of image
processing, by showcasing graphical examples of image processing
operations. Such examples are organized as an array of thumbnails with a
short title (see Fig.~\ref{fig:gallery} left). These thumbnails link to
the webpage of the corresponding example, which features a mini-tutorial
on the image processing method, the code needed to run the example and
the figure generated by the example. Since the graphical gallery is an
efficient way to inform users about the features of
\texttt{scikit-image}, every new feature integrated in the package must
include an example for the gallery.
Longer tutorials and a more narrative documentation is available as well
in the online User Guide of \texttt{scikit-image}. The User Guide explains in
particular "big picture" aspects of \texttt{scikit-image}, such as its
use of \texttt{NumPy} arrays as images, or how the package interacts with
other parts of the scientific Python ecosystem.

@@

Estimating the number of active users of an open-source package is a
difficult task. Statistics of downloads, for example, largely
overestimate the number of active users, all the more if the package is
bundled with others in a software distribution, like Anaconda or
Canopy. A view closer to reality can be obtained by analyzing the
statistics of visits of the online help, available on the project
website.  As of the first half of 2016, 20000 unique visitors visit the
website of scikit-image every month \url{http://scikit-image.org/}. These
users come from 80 countries, located almost equally in Europe, America
and Asia (Oceania and Africa representing a much smaller fraction of
visitors).

@@

\section*{Challenges and roadmap}
